---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# thermokarstdetection

<!-- badges: start -->
<!-- badges: end -->

The package thermokarstdetection provides a function to easily detect thermokarst features from a single digital terrain model (ground surface elevation raster). It identifies cells which are at lower elevation than the surrounding landscape by using a neighborhood function to calculate the average (either median or mean) elevation in one or more user-defined neighborhoods.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("HRodenhizer/thermokarstdetection")
```
## Example

This is a basic example which shows you how to detect thermokarst features using the default options:

```{r example}
library(thermokarstdetection)
library(raster)

### Inspect some elevation data
tk_ponds
plot(tk_ponds[[1]])

### Use tk_detect() to find thermokarst features
tk <- tk_detect(tk_ponds)
```

The main output of `tk_detect()` is a categorical raster of thermokarst class:
```{r thermokarst-output}
# inspect cropped thermokarst output
plot(tk$thermokarst[[1]][[1]])
```

However, it also returns intermediate data products. `elev.crop` is a version of the input elevation raster which has been cropped to the maximum possible output extent of the neighborhood function output:
```{r elevation-output}
plot(tk$elev.crop[[1]])
```

`avg.elev` is the average neighborhood elevation (either median or mean, as specified by the user):
```{r average-elevation-output}
plot(tk$avg.elev[[1]][[1]])
```

`microtopography` is a raster of the deviation in elevation from the average elevation, defined as `elev.crop - avg.elev`:
```{r microtopography-output}
plot(tk$microtopography[[1]][[1]])
```

The output of `tk_detect()` reports the input parameters (radii, fun, cutoff) which were used, as well. `radii` can be either a single value or a vector of radii which determine(s) the neighborhood size when calculation the average elevation. `fun` is the function to use when calculating the neighborhood average elevation (either median or mean). `cutoff` is the value used to classify microtopography as thermokarst.
```{r parameter-output}
tk$radii
tk$fun
tk$cutoff
```
